---
title: "R Notebook"
output: html_notebook
---

## Evidence Stream 2 

Data synthesis for Evidence Stream 2, update to TARA.

Code author: Charlotte Page 

# Install libraries and download data

```{r}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(splitstackshape)
library(stringr)
library(pals)
library(ggsankey)
library(ggalluvial)

es2 <- read.csv("TARA_ES2_extraction_UPDATED.csv")

# Fill all blank cells with NA

es2.na <- es2 %>% mutate_all(na_if,"")

# Filter out studies not included for data extraction 

included <- es2.na  %>% filter(grepl('NA', Exclusion_reason))


```

# Specify a plotting theme for the whole TARA report 

```{r}
my_plot_theme <- function(){
  theme_bw() +
    theme(text = element_text(family = "Helvetica"), # Insert your favourite font
          axis.text = element_text(size = 14), # Y and x axis font size
          axis.title = element_text(size = 16 ), # Axis title font size
          axis.line.x = element_line(color="black"), 
          axis.line.y = element_line(color="black"), 
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),  # panel.grid adds a blank border to your plot.       
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(1, 1, 1, 1), units = , "cm"), # Adjust your plot margins (in cm)
          plot.title = element_text(size = 18, vjust = 1, hjust = 0), # Title font adjustments
          legend.text = element_text(size = 12),          
          legend.title = element_blank(),                              
          legend.position = c(0.95, 0.15), 
          legend.key = element_blank(),
          legend.background = element_rect(color = "black", 
                                           fill = "transparent", 
                                           size = 2, linetype = "blank")) # Sets the legend background to transparent
}

```

# Overall data visualisation 

```{r}

included$Publication_year

# Plot of publication years

a <- ggplot(included, aes(x = Publication_year)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Year") +
        scale_y_continuous(labels = scales::percent, name = "Peer-reviewed articles (%)") +
        #heme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        my_plot_theme() +
  theme_bw() +
  ggtitle("Publication year") 

# Plot of TARA regions 

# Split articles that are associated with multiple regions into seperate rows
included.region.split <- cSplit(included, "TARA_region", ",", "long")

included.region.split$TARA_region <- as.factor(included.region.split$TARA_region)

# To obtain the proportions
included.region.split.table <- table(included.region.split$TARA_region)
prop.table(included.region.split.table)

# To plot the proportions 
b <- ggplot(included.region.split, aes(x =TARA_region)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Region") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Peer-reviewed articles region")


# Plot of TARA marine/estuarine focus 

included$TARA_focus
# Split articles that are associated with multiple regions into seperate rows
included.focus.split <- cSplit(included, "TARA_focus", ",", "long")

included.focus.split$TARA_focus <- as.factor(included.focus.split$TARA_focus)

# To obtain the proportions
focus.table <- table(included.focus.split$TARA_region)
prop.table(focus.table)

# To plot the proportions 
d <- ggplot(included.focus.split, aes(x = TARA_focus)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Focus") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Focus") #+
  #coord_flip()
```


Seperate dataframe by TARA focus 

```{r}

# Coastal and marine 
# Marine 
# If the column contains marine (split by this)
marine <- included %>% filter(grepl('marine', TARA_focus))

# Estuarine 
# If the column contains marine (split by this)
estuarine <- included %>% filter(grepl('Estuarine', TARA_focus))

```


Marine
```{r}

## Overall study location (North, Central, South)

marine$TARA_region

# Split articles that are associated with multiple regions into seperate rows
marine.region.split <- cSplit(marine, "TARA_region", ",", "long")

marine.region.split$TARA_region <- as.factor(marine.region.split$TARA_region)

# To obtain the proportions
marine.region.table <- table(marine.region.split$TARA_region)
prop.table(marine.region.table)

# To plot the proportions 

e <- ggplot(marine.region.split, aes(x = TARA_region)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Region") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine regions") #+
  #coord_flip()


```


Estuarine 

```{r}

## Overall study location (North, Central, South)

estuarine$TARA_region

# Split articles that are associated with multiple regions into seperate rows
estuarine.region.split <- cSplit(estuarine, "TARA_region", ",", "long")

estuarine.region.split$TARA_region <- as.factor(estuarine.region.split$TARA_region)

#estuarine.region.split$TARA_region <- ordered(estuarine.region.split$TARA_region, levels = "	
#1 - North", "2 - Central", "3 - South", "NA")

# To obtain the proportions
estuarine.region.table <- table(marine.region.split$TARA_region)
prop.table(estuarine.region.table)

# To plot the proportions 
f <- ggplot(estuarine.region.split, aes(x = TARA_region)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Region") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine regions") #+
  #coord_flip()


```

# Threat source 

Marine

```{r}

## Filter for threats only (not condition)

# Rename column

marine <- rename(marine, "TARA_Threat_source" = "Threat_source_category_1.Select.more.than.one.of.applicable..See.background.environmental.report.for.threats.that.lie.within.each.of.these.categories..see.page.10.")

str(marine$TARA_Threat_source)

# Filter out articles that only talk about a condition 
# Do this by filtering out NAs in the threat receptor category 

marine.threats <- marine %>% filter(!is.na(TARA_Threat_source))

# Split articles that are associated with multiple regions into seperate rows
marine.threat.split <- cSplit(marine.threats, "TARA_Threat_source", ",", "long")

marine.threat.split$TARA_Threat_source <- as.factor(marine.threat.split$TARA_Threat_source)

# To obtain the proportions
marine.threat.table <- table(marine.threat.split$TARA_Threat_source)
prop.table(marine.threat.table)

# To plot the proportions 
ggplot(marine.threat.split, aes(x = TARA_Threat_source)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Threat category") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  ggtitle("Marine threats") +
  theme_bw() +
  coord_flip()



```

Estuarine 

```{r}
## Overall study location (North, Central, South)

# Rename column
estuarine <- rename(estuarine, "TARA_Threat_source" = "Threat_source_category_1.Select.more.than.one.of.applicable..See.background.environmental.report.for.threats.that.lie.within.each.of.these.categories..see.page.10.")

str(estuarine$TARA_Threat_source)

# Filter out articles that only talk about a condition 
# Do this by filtering out NAs in the threat receptor category 

estuarine.threats <- estuarine %>% filter(!is.na(TARA_Threat_source))

# Split articles that are associated with multiple regions into seperate rows
estuarine.threat.split <- cSplit(estuarine.threats, "TARA_Threat_source", ",", "long")

estuarine.threat.split$TARA_Threat_source <- as.factor(estuarine.threat.split$TARA_Threat_source)

# To obtain the proportions
estuarine.threat.table <- table(estuarine.threat.split$TARA_Threat_source)
prop.table(estuarine.threat.table)

# To plot the proportions 
ggplot(estuarine.threat.split, aes(x = TARA_Threat_source)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Threat category") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  ggtitle("Estuarine threats") +
  theme_bw() +
  coord_flip()

```
Do not need to plot these (just land based activities for all)

Land based activities 

```{r}

##### Resouce use activities 

##### Split articles that are associated with multiple threats into seperate rows
marine.threatcat.split <- cSplit(marine.threats, "Threat_source", ";", "long")

marine.threatcat.split$Threat_source <- as.factor(marine.threatcat.split$Threat_source)

mar.land.use <- marine.threatcat.split %>% filter(TARA_Threat_source == "Resource use")

# To plot the proportions 

mar.lu.threat.cat.proportion <- mar.land.use  %>% group_by(Threat_source) %>% summarise(prop = (Percentage=n()/nrow(.))*100)

# Truncate the names 

mar.lu.threat.cat.proportion.shrt <- apply(mar.lu.threat.cat.proportion, 2, strtrim,3)
mar.lu.threat.cat.proportion.frame <- as.data.frame(mar.lu.threat.cat.proportion.shrt )

# Compute the cumulative percentages (top of each rectangle)
mar.lu.threat.cat.proportion.frame$ymax = cumsum(mar.lu.threat.cat.proportion.frame$prop)

# Compute the bottom of each rectangle
mar.lu.threat.cat.proportion.frame$ymin = c(0, head(mar.lu.threat.cat.proportion.frame$ymax, n=-1))

# Compute label position
mar.lu.threat.cat.proportion.frame$labelPosition <- (mar.lu.threat.cat.proportion.frame$ymax + mar.lu.threat.cat.proportion.frame$ymin) / 2

# Compute a good label
mar.lu.threat.cat.proportion.frame$label <- paste0(mar.lu.threat.cat.proportion.frame$Threat_source, "\n value: ", mar.lu.threat.cat.proportion.frame$prop)

# Make the plot
ggplot(mar.lu.threat.cat.proportion.frame, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Threat_source)) +
  geom_rect() +
  geom_text( x=5, aes(y=labelPosition, label=label, color=Threat_source), size=4) + # x here controls label position (inner / outer)
 scale_fill_brewer(palette="PRGn") +
scale_color_brewer(palette="PRGn") +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() 
 # theme(legend.position = "none")

```

```{r}

##### Resource use activities 

# 14 colour paletter needed.

##### Split articles that are associated with multiple threats into seperate rows
estuarine.threatcat.split <- cSplit(estuarine.threats, "Threat_source", ";", "long")

estuarine.threatcat.split$Threat_source <- as.factor(estuarine.threatcat.split$Threat_source)

est.land.use <- estuarine.threatcat.split %>% filter(TARA_Threat_source == "Land based activities")

# To plot the proportions 

est.lu.threat.cat.proportion <- est.land.use  %>% group_by(Threat_source) %>% summarise(prop = (Percentage=n()/nrow(.))*100)

# Truncate the names 

est.lu.threat.cat.proportion.shrt <- apply(est.lu.threat.cat.proportion, 2, strtrim,3)
est.lu.threat.cat.proportion.frame <- as.data.frame(est.lu.threat.cat.proportion.shrt )

# Compute the cumulative percentages (top of each rectangle)
est.lu.threat.cat.proportion.frame$ymax = cumsum(est.lu.threat.cat.proportion.frame$prop)

# Compute the bottom of each rectangle
est.lu.threat.cat.proportion.frame$ymin = c(0, head(est.lu.threat.cat.proportion.frame$ymax, n=-1))

# Compute label position
est.lu.threat.cat.proportion.frame$labelPosition <- (est.lu.threat.cat.proportion.frame$ymax + est.lu.threat.cat.proportion.frame$ymin) / 2

# Compute a good label
est.lu.threat.cat.proportion.frame$label <- paste0(est.lu.threat.cat.proportion.frame$Threat_source, "\n value: ", est.lu.threat.cat.proportion.frame$prop)

ggplot(est.lu.threat.cat.proportion.frame, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Threat_source)) +
  geom_rect() +
  geom_text( x=5, aes(y=labelPosition, label=label, color=Threat_source), size=4) + # x here controls label position (inner / outer)
 scale_fill_manual(values=as.vector(stepped(11))) +
scale_colour_manual(values=as.vector(stepped(11))) +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() 
 # theme(legend.position = "none")


```

Threat receptor relationships 

# Marine assets 

```{r}

##### Split articles that are associated with multiple threats into separate rows
# Split articles that are associated with multiple threats into seperate rows
marine.threat.receptor <- cSplit(marine.threatcat.split, "Receptor_category.threat_receptor", ";", "long")


marine.threat.receptor$Receptor_category.threat_receptor <- as.factor(marine.threat.receptor$Receptor_category.threat_receptor)


# Resource use
mar.lu.tr <- marine.threat.receptor %>% filter(TARA_Threat_source == "Resource use")
mar.lu <- select(mar.lu.tr,Threat_source,
       Specific_threat_source..If.detailed.specify.the.actual.threat.activity..If.multiple.threat.activities.are.listed.seperate.terms.in.a.list.with..,Receptor_category.threat_receptor,
       Specific_receptor_category.threat_receptor.If.detailed.list.the.specific.receptor..If.there.is.more.than.one.receptor.measured.list.these.as.terms.separated.by..,Comment.for.extraction)

write.csv(mar.lu,"ES2_Marine_threat_receptors_land_based_activities.csv")

```


```{r}
mar.lu.tr

ggplot(data = mar.lu.tr ,
       aes(axis1 = Threat_source, axis2 = Receptor_category.threat_receptor)) +
  geom_alluvium(aes(fill = Receptor_category.threat_receptor)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void() +
     scale_fill_brewer(palette="Spectral") +
 scale_color_brewer(palette="Spectral") +
  theme(legend.position = "none")

```

# Estuarine assets 

```{r}

##### Split articles that are associated with multiple threats into separate rows
# Split articles that are associated with multiple threats into seperate rows
estuarine.threat.receptor <- cSplit(estuarine.threatcat.split, "Receptor_category.threat_receptor", ";", "long")


estuarine.threat.receptor$Receptor_category.threat_receptor <- as.factor(estuarine.threat.receptor$Receptor_category.threat_receptor)

# Land based activities
est.lu.tr <- estuarine.threat.receptor %>% filter(TARA_Threat_source == "Land based activities")

est.lu <- select(est.lu.tr,Threat_source,
       Specific_threat_source..If.detailed.specify.the.actual.threat.activity..If.multiple.threat.activities.are.listed.seperate.terms.in.a.list.with..,Receptor_category.threat_receptor,
       Specific_receptor_category.threat_receptor.If.detailed.list.the.specific.receptor..If.there.is.more.than.one.receptor.measured.list.these.as.terms.separated.by..,Comment.for.extraction)

write.csv(est.lu,"ES2_Estuarine_threat_receptors_land_based_activities.csv")

```

```{r}

est.lu.tr

ggplot(data = est.lu.tr ,
       aes(axis1 = Threat_source, axis2 = Receptor_category.threat_receptor)) +
  geom_alluvium(aes(fill = Receptor_category.threat_receptor)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void() +
          scale_fill_brewer(palette="Spectral") +
 scale_color_brewer(palette="Spectral") +
  theme(legend.position = "none")

# save as a csv for appendix 

write.csv(est.lu.tr, "ES2_Estuary_land_based_activities_refs.csv")
```

Asset conditions 

# Condition of marine asset

```{r}

marine <- rename(marine, "TARA_asset_condition" = "Condition_trend.If.the.condition.of.the.asset.is.discussed.specify.the.outcome.of.the.asset.compared.to.the.comparator..")

# Remove NAs from condition trend 

# Filter out studies not included for data extraction 

marine_condition <- marine %>% filter(!is.na(TARA_asset_condition))

write.csv(marine_condition,"ES2_Coastalandmarine_assets_condition_refs.csv")

marine_condition.split.location <- cSplit(marine_condition, "TARA_region", ",", "long")

# Location of conditions assets 

marine_condition.split.location$TARA_region <- as.factor(marine_condition.split.location$TARA_region)

# To obtain the proportions
marine_condition.table <- table(marine_condition.split.location$TARA_region)
prop.table(marine_condition.table)

# To plot the proportions 

ggplot(marine_condition.split.location, aes(x = TARA_region)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Region") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine regions") #+
  #coord_flip()


# Trend of condition 


marine_condition.split.asset <- cSplit(marine_condition,"Receptor_category.threat_receptor", ";", "long")

test <- marine_condition.split.asset   %>% group_by(Receptor_category.threat_receptor,TARA_asset_condition) %>% summarise(prop = (Percentage=n()/nrow(.))*100)

view(test)

# Truncate the names 

test.a <- apply(test, 2, strtrim,3)
test.a.frame <- as.data.frame(test.a)

ggplot(test.a.frame, aes(x= Receptor_category.threat_receptor, y = prop, fill = TARA_asset_condition)) + 
  geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
  xlab("Marine asset") +
  my_plot_theme() +
  theme_bw() +
  ggtitle("Marine asset condition trends") +
  coord_flip() +
  scale_fill_brewer(palette="Set1") +
 scale_color_brewer(palette="Set1") 

```





Estuarine 


```{r}

estuarine <- rename(estuarine, "TARA_asset_condition" = "Condition_trend.If.the.condition.of.the.asset.is.discussed.specify.the.outcome.of.the.asset.compared.to.the.comparator..")

# Remove NAs from condition trend 

# Filter out studies not included for data extraction 

estuarine_condition <- estuarine %>% filter(!is.na(TARA_asset_condition))

write.csv(estuarine_condition,"Estuarine_assets_condition_refs.csv")

estuarine_condition.split.location <- cSplit(estuarine_condition, "TARA_region", ",", "long")

# Location of conditions assets 

estuarine_condition.split.location$TARA_region <- as.factor(estuarine_condition.split.location$TARA_region)

# To obtain the proportions
estuarine_condition.table <- table(estuarine_condition.split.location$TARA_region)
prop.table(estuarine_condition.table)

# To plot the proportions 

ggplot(estuarine_condition.split.location, aes(x = TARA_region)) +
        geom_bar(aes(y = (..count..)/sum(..count..))) +
        xlab("Region") +
        scale_y_continuous(labels = scales::percent, name = "Proportion") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine regions") #+
  #coord_flip()


# Trend of condition 

estuarine_condition.split.asset <- cSplit(estuarine_condition,"Receptor_category.threat_receptor", ";", "long")

View(estuarine_condition.split.asset)
test.est <- estuarine_condition.split.asset   %>% group_by(Receptor_category.threat_receptor,TARA_asset_condition) %>% summarise(prop = (Percentage=n()/nrow(.))*100)

view(test.est)

# Truncate the names 

test.b <- apply(test.est, 2, strtrim,3)
test.b.frame <- as.data.frame(test.b)

#write.csv(test.b.frame, "Estuary_conditionplot_edit.csv")
test.b.edit <- read.csv("Estuary_conditionplot_edit.csv")

# Redorder levels of variable

test.b.edit$TARA_asset_condition <- factor(test.b.edit$TARA_asset_condition, levels = c("Neg", "Pos", "Oth","Neu"))

ggplot(test.b.edit, aes(x= Receptor_category.threat_receptor, y = prop, fill = TARA_asset_condition)) + 
 geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
          xlab("Estuarine asset") +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine asset condition trends") +
    coord_flip() +
  scale_fill_brewer(palette="Set1") +
 scale_color_brewer(palette="Set1") 

```


