---
title: "R Notebook"
output: html_notebook
---

Knowledge-gap filling 


```{r}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(splitstackshape)
library(stringr)
library(pals)
library(ggsankey)
library(ggalluvial)
library(paletteer)

# EVIDENCE STREAM 1

es1 <- read.csv("TARA_ES1_data_extraction.csv")

# Fill all blank cells with NA

es1.na <- es1 %>% mutate_all(na_if,"")

# Filter out studies not included for data extraction 

es.1_included <- es1.na  %>% filter(grepl('NA', Exclusion_reason))

es.1_included$Receptor_category.threat_receptor

es1_dataframe <- es.1_included %>%
  select(TARA_focus,TARA_region,
         Threat_source_category_1.Select.more.than.one.of.applicable.,
         Threat_source,
         Receptor_category.threat_receptor) %>%
  mutate(Evidence_stream = "ES_1") %>%
  rename(Threat_source_category=Threat_source_category_1.Select.more.than.one.of.applicable.)





```

```{r}
es2 <- read.csv("TARA_ES2_extraction_updated.csv")

# Fill all blank cells with NA

es2.na <- es2 %>% mutate_all(na_if,"")

# Filter out studies not included for data extraction 

es.2_included <- es2.na  %>% filter(grepl('NA', Exclusion_reason))

es2_dataframe <- es.2_included %>%
  select(TARA_focus,TARA_region,
         Threat_source_category_1.Select.more.than.one.of.applicable..See.background.environmental.report.for.threats.that.lie.within.each.of.these.categories..see.page.10.,
         Threat_source,
         Receptor_category.threat_receptor) %>%
  mutate(Evidence_stream = "ES_2") %>%
  rename(Threat_source_category = Threat_source_category_1.Select.more.than.one.of.applicable..See.background.environmental.report.for.threats.that.lie.within.each.of.these.categories..see.page.10.)

```

# Specify a plotting theme for the whole TARA report 

```{r}
my_plot_theme <- function(){
  theme_bw() +
    theme(text = element_text(family = "Helvetica"), # Insert your favourite font
          axis.text = element_text(size = 14), # Y and x axis font size
          axis.title = element_text(size = 16 ), # Axis title font size
          axis.line.x = element_line(color="black"), 
          axis.line.y = element_line(color="black"), 
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),  # panel.grid adds a blank border to your plot.       
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(1, 1, 1, 1), units = , "cm"), # Adjust your plot margins (in cm)
          plot.title = element_text(size = 18, vjust = 1, hjust = 0), # Title font adjustments
          legend.text = element_text(size = 12),          
          legend.title = element_blank(),                              
          legend.position = c(0.95, 0.15), 
          legend.key = element_blank(),
          legend.background = element_rect(color = "black", 
                                           fill = "transparent", 
                                           size = 2, linetype = "blank")) # Sets the legend background to transparent
}

```

Join the data frames together 

```{r}

es <- rbind(es1_dataframe,es2_dataframe)

```

# Split by marine and estuarine 

```{r}

# Coastal and marine 
# Marine 
# If the column contains marine (split by this)
marine <- es %>% filter(grepl('marine', TARA_focus))

# Estuarine 
# If the column contains marine (split by this)
estuarine <- es %>% filter(grepl('Estuarine', TARA_focus))

```

Overall study locations covered by both ES 

```{r}
## Overall study location (North, Central, South)

marine$TARA_region

# Split articles that are associated with multiple regions into seperate rows
marine.region.split <- cSplit(marine, "TARA_region", ",", "long")

marine.region.split$TARA_region <- as.factor(marine.region.split$TARA_region)
marine.region.split$Evidence_stream <- as.factor(marine.region.split$Evidence_stream)

# To plot the proportions 

ggplot(marine.region.split, aes(x = TARA_region)) +
        geom_bar() +
        xlab("Region") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine regions") + facet_wrap(~Evidence_stream)


ggplot(marine.region.split, aes(x = TARA_region, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Region") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine regions") 


```


```{r}
## Overall study location (North, Central, South)

estuarine$TARA_region

# Split articles that are associated with multiple regions into seperate rows
estuarine.region.split <- cSplit(estuarine, "TARA_region", ",", "long")

estuarine.region.split$TARA_region <- as.factor(estuarine.region.split$TARA_region)
estuarine.region.split$Evidence_stream <- as.factor(estuarine.region.split$Evidence_stream)

# To plot the proportions 

ggplot(estuarine.region.split, aes(x = TARA_region)) +
        geom_bar() +
        xlab("Region") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine regions") + facet_wrap(~Evidence_stream)


ggplot(estuarine.region.split, aes(x = TARA_region, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Region") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine regions") 
```
# Threat categories 

```{r}

marine$Threat_source_category

# Split articles that are associated with multiple regions into seperate rows
marine.threat.split <- cSplit(marine, "Threat_source_category", ",", "long")

marine.threat.split $TARA_region <- as.factor(marine.threat.split $TARA_region)
marine.threat.split $Evidence_stream <- as.factor(marine.threat.split$Evidence_stream)

# Filter out NAs 
marine.threat.split <- marine.threat.split %>% filter(!is.na(Threat_source_category))

# To plot the proportions 

ggplot(marine.threat.split , aes(x = Threat_source_category)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine threats") + facet_wrap(~Evidence_stream)


ggplot(marine.threat.split , aes(x = Threat_source_category, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine threats") 

```
```{r}

estuarine$Threat_source_category

# Split articles that are associated with multiple regions into seperate rows
estuarine.threat.split <- cSplit(estuarine, "Threat_source_category", ",", "long")

estuarine.threat.split $TARA_region <- as.factor(estuarine.threat.split $TARA_region)
estuarine.threat.split $Evidence_stream <- as.factor(estuarine.threat.split$Evidence_stream)

# Filter out NAs 
estuarine.threat.split <- estuarine.threat.split %>% filter(!is.na(Threat_source_category))

# To plot the proportions 

ggplot(estuarine.threat.split , aes(x = Threat_source_category)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine threats") + facet_wrap(~Evidence_stream)


ggplot(estuarine.threat.split , aes(x = Threat_source_category, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Estuarine threats") 

```

## Threats 

```{r}

##### Split articles that are associated with multiple threats into seperate rows
marine.threatcat.split <- cSplit(marine, "Threat_source", ";", "long")

marine.threatcat.split$Threat_source <- as.factor(marine.threatcat.split$Threat_source)

# Filter out NAs 
marine.threatcat.split <- marine.threatcat.split %>% filter(!is.na(Threat_source))


write.csv(marine.threatcat.split,"E3_marine_threats.csv")

########

marinethreats <- read.csv("E3_marine_threats_updated.csv") 


# Filter for climate change 

ccmarinegaps <- marinethreats %>%
  filter(Threat_source_category == "Climate change") %>%
  select(Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,4)

ccmarinegaps.frame <- as.data.frame(ccmarinegaps) 

ggplot(ccmarinegaps.frame, aes(x = Threat_source, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine threats - Climate change") +
  coord_flip() 

# Filter for resource use 

rumarinegaps <- marinethreats %>%
  filter(Threat_source_category == "Resource use") %>%
  select(Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,4)

rumarinegaps.frame <- as.data.frame(rumarinegaps) 

ggplot(rumarinegaps.frame, aes(x = Threat_source, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine threats - Resource use") +
  coord_flip() 

# Filter for landbased activities 

lbamarinegaps <- marinethreats %>%
  filter(Threat_source_category == "Land based activities") %>%
  select(Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,4)

lbamarinegaps.frame <- as.data.frame(lbamarinegaps) 

ggplot(lbamarinegaps.frame, aes(x = Threat_source, fill = Evidence_stream)) +
        geom_bar() +
        xlab("Threat category") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    my_plot_theme() +
  theme_bw() +
  ggtitle("Marine threats - Land based activities") +
  coord_flip() 

```

Threats 

```{r}

##### Split articles that are associated with multiple threats into seperate rows
estuarine.threatcat.split <- cSplit(estuarine, "Threat_source", ";", "long")

estuarine.threatcat.split$Threat_source <- as.factor(estuarine.threatcat.split$Threat_source)

# Filter out NAs 
estuarine.threatcat.split <- estuarine.threatcat.split %>% filter(!is.na(Threat_source))


write.csv(estuarine.threatcat.split,"E3_estuarine_threats.csv")

estuarinethreats <- read.csv("E3_estuarine_threats_updated.csv") 

```

```{r}
# Climate change 

ccmarinegaps <- marinethreats %>%
  filter(Threat_source_category == "Climate change") %>%
  select(TARA_focus,Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,20)

ccestuarinegaps <- estuarinethreats %>%
  filter(Threat_source_category == "Climate change") %>%
  select(TARA_focus,Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,20)

ccgaps <- rbind(ccmarinegaps,ccestuarinegaps)

str(ccgaps)

ccgaps <- as.data.frame(ccgaps) 

ccgaps %>% 
count(Threat_source,TARA_focus,.drop=FALSE) %>% 
ggplot(aes(x=TARA_focus,y=Threat_source,fill=n)) + geom_tile() + theme_classic() +scale_fill_distiller(palette = "YlOrRd", direction = 1)

```

```{r}
# Resource use

rumarinegaps <- marinethreats %>%
  filter(Threat_source_category == "Resource use") %>%
  select(TARA_focus,Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,20)


ruestuarinegaps <- estuarinethreats %>%
  filter(Threat_source_category == "Resource use") %>%
  select(TARA_focus,Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,20)

rugaps <- rbind(rumarinegaps,ruestuarinegaps)

rugaps <- as.data.frame(rugaps) 

View(rugaps)

rugaps %>% 
count(Threat_source,TARA_focus,.drop=FALSE) %>% 
ggplot(aes(x=TARA_focus,y=Threat_source,fill=n)) + geom_tile() + theme_classic() +scale_fill_distiller(palette = "YlOrRd", direction = 1)

```

```{r}
# Land based activities

lumarinegaps <- marinethreats %>%
  filter(Threat_source_category == "Land based activities") %>%
  select(TARA_focus,Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,20)


luestuarinegaps <- estuarinethreats %>%
  filter(Threat_source_category == "Land based activities") %>%
  select(TARA_focus,Threat_source_category,Threat_source,Evidence_stream) %>%
  apply(2, strtrim,20)

lugaps <- rbind(lumarinegaps,luestuarinegaps)

lugaps <- as.data.frame(lugaps) 

View(lugaps)

lugaps %>% 
count(Threat_source,TARA_focus,.drop=FALSE) %>% 
ggplot(aes(x=TARA_focus,y=Threat_source,fill=n)) + geom_tile() + theme_classic() +scale_fill_distiller(palette = "YlOrRd", direction = 1)

```


## Threats and their impacts on marine assets

```{r}
##### Split articles that are associated with multiple threats into seperate rows

marine$Receptor_category.threat_receptor
marine.receptor <- cSplit(marine, "Receptor_category.threat_receptor", ";", "long")

marine.receptor <- marine.receptor %>% filter(!is.na(Receptor_category.threat_receptor))

# Filter out just marine/ocean
wrtie.csv(marine.receptor,"")


marine.threatcat.split$Threat_source <- as.factor(marine.threatcat.split$Threat_source)

# Truncate the names 

marine.threatcat.split.shrt <- apply(marine.threatcat.split, 3, strtrim,4)
marine.threatcat.split.frame <- as.data.frame(marine.threatcat.split.shrt)

ggplot(data = marine.receptor  ,
       aes(axis1 = Evidence_stream, axis2 = Receptor_category.threat_receptor)) +
  geom_alluvium(aes(fill = Evidence_stream)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void() +
     scale_fill_brewer(palette="Spectral") +
 scale_color_brewer(palette="Spectral") +
  theme(legend.position = "none")
```

## Make a sankey diagram for ES1 and ES2 combined

```{r}

str(es)

##### Split articles that are associated with multiple threats into separate rows
# Split articles that are associated with multiple threats into seperate rows
es.threatreceptor <- cSplit(es, "Threat_source", ";", "long")

# Split articles that are associated with multiple threat categories into seperate rows
es.threatreceptorcat <- cSplit(es.threatreceptor, "Threat_source_category", ",", "long")

# Split articles that are associated with multiple threat receptors into seperate rows
es.threatreceptor.cat <- cSplit(es.threatreceptorcat, "Receptor_category.threat_receptor", ";", "long")

# Split articles that are associated with TARA focus into seperate rows
es.threatreceptor.cat.focus <- cSplit(es.threatreceptor.cat, "TARA_focus", ",", "long")

View(es.threatreceptor.cat.focus)

# Filter for Coastal & Marine, resource use 

es1es2marine <- es.threatreceptor.cat.focus %>% filter(Threat_source_category == "Resource use") %>% 
  filter(TARA_focus == "1 - Coastal and marine")

# Filter for Estuarine, land based activities

es1es2estuarine <- es.threatreceptor.cat.focus %>% filter(Threat_source_category == "Land based activities") %>% 
  filter(TARA_focus == "2 - Estuarine")

```


Sankey diagrams for: 

Coastal and marine, resource use


```{r}


ggplot(data = es1es2marine,
       aes(axis1 = Threat_source, axis2 = Receptor_category.threat_receptor)) +
  geom_alluvium(aes(fill = Receptor_category.threat_receptor)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void() +
   scale_fill_brewer(palette="Spectral") +
 scale_color_brewer(palette="Spectral") +
  theme(legend.position = "none")


```

Donut plot 

```{r}


es1es2marine.proportion <- es1es2marine %>% group_by(Threat_source) %>% summarise(prop = (Percentage=n()/nrow(.))*100)

# Truncate the names 

es1es2marine.proportion.shrt <- apply(es1es2marine.proportion, 2, strtrim,3)
es1es2marine.proportion.shrt.frame <- as.data.frame(es1es2marine.proportion.shrt)

# Compute the cumulative percentages (top of each rectangle)
es1es2marine.proportion.shrt.frame$ymax = cumsum(es1es2marine.proportion.shrt.frame$prop)

# Compute the bottom of each rectangle
es1es2marine.proportion.shrt.frame$ymin = c(0, head(es1es2marine.proportion.shrt.frame$ymax, n=-1))

# Compute label position
es1es2marine.proportion.shrt.frame$labelPosition <- (es1es2marine.proportion.shrt.frame$ymax + es1es2marine.proportion.shrt.frame$ymin) / 2

# Compute a good label
es1es2marine.proportion.shrt.frame$label <- paste0(es1es2marine.proportion.shrt.frame$Threat_source, "\n value: ", es1es2marine.proportion.shrt.frame$prop)

# Make the plot
ggplot(es1es2marine.proportion.shrt.frame, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Threat_source)) +
  geom_rect() +
  geom_text( x=5, aes(y=labelPosition, label=label, color=Threat_source), size=4) + # x here controls label position (inner / outer)
 scale_fill_brewer(palette="Paired") +
 scale_color_brewer(palette="Paired") +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() 
 # theme(legend.position = "none")


```


Estuarine, landbased activities


```{r}

ggplot(data = es1es2estuarine,
       aes(axis1 = Threat_source, axis2 = Receptor_category.threat_receptor)) +
  geom_alluvium(aes(fill = Receptor_category.threat_receptor)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void() +
   #scale_fill_brewer(palette="Spectral") +
 #scale_color_brewer(palette="Spectral") +
  theme(legend.position = "none")

```

Donut plot

```{r}

es1es2estuarine.proportion <- es1es2estuarine %>% group_by(Threat_source) %>% summarise(prop = (Percentage=n()/nrow(.))*100)

# Truncate the names 

es1es2estuarine.proportion.shrt <- apply(es1es2estuarine.proportion, 2, strtrim,3)
es1es2estuarine.proportion.shrt.frame <- as.data.frame(es1es2estuarine.proportion.shrt)

# Compute the cumulative percentages (top of each rectangle)
es1es2estuarine.proportion.shrt.frame$ymax = cumsum(es1es2estuarine.proportion.shrt.frame$prop)

# Compute the bottom of each rectangle
es1es2estuarine.proportion.shrt.frame$ymin = c(0, head(es1es2estuarine.proportion.shrt.frame$ymax, n=-1))

# Compute label position
es1es2estuarine.proportion.shrt.frame$labelPosition <- (es1es2estuarine.proportion.shrt.frame$ymax + es1es2estuarine.proportion.shrt.frame$ymin) / 2

# Compute a good label
es1es2estuarine.proportion.shrt.frame$label <- paste0(es1es2estuarine.proportion.shrt.frame$Threat_source, "\n value: ", es1es2estuarine.proportion.shrt.frame$prop)

# Make the plot
ggplot(es1es2estuarine.proportion.shrt.frame, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Threat_source)) +
  geom_rect() +
  geom_text( x=5, aes(y=labelPosition, label=label, color=Threat_source), size=4) + # x here controls label position (inner / outer)
 scale_fill_paletteer_d("ggthemes::Classic_20") +
 scale_color_paletteer_d("ggthemes::Classic_20") +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() 
 # theme(legend.position = "none")





```



